# Negative Binomial Hurdle Models with Zero-Heavy Count Data {.unnumbered}
## Modelling Alcohol and Other Drug Use, Psychological Distress, and Occupational Functioning in Australian Public Safety Personnel

(full script at bottom of page)

## Research Question/Aims
1. Examine rates of severity of alcohol and other drug (AOD) use among public safety personnel (PSP)
2. Examine associations between AOD use, psychological distress, and occupational functioning (absenteeism and presenteeism) among PSP
3. **Examine whether AOD use and psychological distress predict impairment to occupational function, both independently, and in combination **

## Overview 
Examining how alcohol and other drug use and psychological distress relate to absenteeism and presenteeism in public safety personnel. 

Key Variables:

- Psychological distress (K10) 
- Substance Use (ASSIST) 
- Workplace Outcomes

Outcomes:

- Absenteeism and Presenteeism (from k10+)
- Count data with many zeros 

```{r set, include = FALSE}
pacman::p_load(here, tidyverse, skimr, gtsummary, flextable, officer,
       scales, rstatix, GGally, janitor, DataExplorer, sjmisc,
       vtable, psych, car, lmtest, naniar, 
       GGally, ggpubr, margins, cowplot,
       MASS, pscl, AER, performance, apaTables, Hmisc, glmmTMB, margins, dplyr, SmartEDA, summarytools,
       ggstatsplot, rcompanion, dunn.test)

here::i_am("posts/icare.qmd")
df <- readRDS(here::here("R files/clean_self_assessment.RData"))
df <- df %>% filter(assessment_time == "Time 1")
```

## Data Pre-processing/Screening 
Initial visualisation of absenteeism and presenteeism showed:

- strong right-skew 
- large proportion of zeros 
```{r plot1, warning = FALSE, message = FALSE}
df %>% 
  dplyr::select(
    Absenteeism = k10_11_absenteeism, Presenteeism = k10_12_presenteeism) |> 
  pivot_longer(
    cols = c(Absenteeism, Presenteeism),
    names_to = "Variable",
    values_to = "Days") |> 
    filter(!is.na(Days)) |> 
    gghistogram(
        x = "Days", 
        y = "count", 
        xlab = "Days", 
        ylab = FALSE, 
        add = "median",
        facet.by = "Variable", 
        binwidth = 1
    ) +
    scale_x_continuous(breaks = seq(0, 30, by = 2))
```

**note: suggests standard linear models would be inappropriate. count-based better** 

## Model selection 
Given the distributional properties of absenteeism and presenteeism, I considered several candidate models:

- Poisson regression
- Negative binomial regression 
- Zero-inflated Poisson (ZIP)
- Zero-inflated negative binomial models (ZINB) 
- Hurdle negative binomial models 

### Considerations: 
**Overdispersion**: Poisson models assume that the mean = variance. Preliminary checks indicated overdispersion, making Poisson models unsuitable. 

**Excess zeros**: The data contained more zeros than expected under standard count models. This raises 2 choices:

- Zero inflated models: assume two latent groups (always-zero vs always at-risk) 
- Hurdle models: assume a two-stage process (zero vs positive counts) 

## Modelling strategy 
### Univariate models - all types 
**Here I made a function that ran a series of univariate models on each outcome to check dispersion**
This helped to evaluate which count model handled the outcome data the best. 
Univariate predictors included age, gender, years of employment, k10 total score, maximum alcohol use risk score, and maximum other drug use risk score. 

Models were compared using: AIC/BIC (fit), dispersion stats, vuong tests, predicted vs observed zeros.

Although zero-inflated models and hurdle models were basically the same, MS and I discussed the choice from a conceptual angle, and we agreed that the zero-inflation was too rigid with the "always-zero" or "always at-risk" approach. 

Given we are looking at **absenteeism** and **presenteeism**, days of each is unlikely to be **always**. 

So we chose **Hurdle Negative Binomial Models.**

#### Reasons: 
- they count for overdispersion 
- explicitly model the probability of any absenteeism/presenteeism (logistic) and the frequency of absenteeism/presenteeism among those who didn't have zero values (count) 
- conceptually, they distinguish between whether someone is absent and how often they are absent. 
- more aligned with our aims/RQs 

Once Hurdle Models were chosen, each predictor was first examined in isolation to:

- identify potential associations
- understand variable-specific effects 
- provide context before multivariable analyses 



