# Project Setup {.unnumbered}
## purpose
- keep everything organised 
- reproducible and transparent structure 
- clear workflow/file organisation 

## structure 
- create project root folder
- open in software 
```
project/

├── R/
├── data/
│   ├── raw/        # source data (never edited)
│   ├── versions/   # timestamped snapshots
│   ├── processed/  # cleaned / merged
│   └── final/      # datasets used for analysis / sharing
├── outputs/        # figures, tables, model objects
├── docs/           # protocols, codebooks, notes
├── renv.lock
```
**notes**:
- root folder always treated as root 
- no `setwd()`
- all paths relative 
- never change raw data 

## environment 
initialise on first open: 
```{r, eval = FALSE}
# create renv and lockfile in root folder 
renv::init()

# install packages 
install.packages(c("tidyverse", "here", "renv", "usethis", "devtools"))

# snapshot to save in lockfile for reproducibility and collaboration
renv::snapshot()
```

**notes**: 
- `renv.lockfile` tracks package versions 
- collaborators use `renv::restore()` to get same environment  

## paths 
- anchor scripts using `here` package 
- avoid using absolute paths 
```{r, run = FALSE, eval = FALSE}
# shows where file is saved relative to root 
here::i_am("R/01_data_processing.R")

# import data using relative paths
readr::read_csv(here::here("data/raw/survey_data_raw.csv"))
```

## secrets 
- use `usethis` to create `.Renviron` file in root folder and add secrets in there 
- for example: API keys, database credentials etc. 
- github will ignore this file so its not shared with collaborators 

```{r, run = FALSE, eval = FALSE} 
usethis::edit_r_environ()
```
- after running, edit `.Renviron` file 
- format specifically: 
```
API_KEY="your_api_key_here"
DB_USERNAME="your_db_username_here"
DB_PASSWORD="your_db_password_here"
```
after saving the `.Renviron` file, access the secrets in your R code:
```{r, run=FALSE, evaluate=FALSE}
api_key <- Sys.getenv("API_KEY")
db_username <- Sys.getenv("DB_USERNAME")
db_password <- Sys.getenv("DB_PASSWORD")
```
