# Script Practices {.unnumbered}
- clear and organised scripts
- modular code that is easy to read and maintain
- clear workflow and order of operations

## organisation
- one script = one purpose 
- avoid combining stages of large analyses 
- keep scripts short 

Example layout of script folder:
```
R/
├── 01_data_processing.R   # scripts for cleaning and processing data
├── 02_linear_models.R     # scripts for running linear models 
├── utilities.R           # scripts for functions that are used across multiple analyses
├── functions_data_processing.R  # scripts for functions that are specific to data processing
```

## functions 
- any repeated logic gets moved to functions 
- if repeated across different tasks -> `utilities.R`
- if task-specific -> `functions_01_data_cleaning.R`
- source those scripts when needed in analyses 

## style conventions 
- Use pipes: `|>`
- one step per line 

```{r, eval=FALSE, run=FALSE}
df |>
    dplyr::filter(age > 18) |>  
    dplyr::mutate(age_group = ifelse(age < 30, "young", "old")) |>
    dplyr::group_by(age_group) |>
    dplyr::summarise(mean_income = mean(income))
```

## naming conventions 
- use `snake_case` for everything
- clear descriptive names (not `df1`)
- objects start with noun (e.g., `df_*` or `mod_*`)
- functions start with verb (e.g., `clean_`, `fit_`)

## ordering and flow 
- if structure begins to exist (e.g., order of scripts) use `targets`
- removes manual ordering and execution 
- tracks dependencies 
```{r, eval = FALSE, run = FALSE}
list(
  tar_target(raw, get_data()),
  tar_target(cleaned, clean_data(raw)),
  tar_target(model, fit_linear_model(cleaned)),
  tar_target(plot, create_plot(model))
)
```
